---
title: "ProPublica Facebook Ad Collector"
date: "Last updated: `r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(dplyr)
library(lubridate)
library(tidyr)
library(purrr)
library(xray)
library(feather)
library(ggplot2)
library(scales)
library(htmlwidgets)
library(ochRe)
library(dutchmasters)
set.seed(2018)

source("../inst/load_propub_fb.R")

theme_set(theme_grey() + 
            theme(panel.background = element_rect(fil = dutchmasters$little_street[[2]]),
                  panel.grid = element_blank(),
                  panel.grid.major.y = element_line(colour = dutchmasters$little_street[[11]])))
```

```{r sample_ad_db}
ads_sample <- ads_feather() %>% 
  sample_frac(0.05) %>% 
  mutate_at(vars(matches("(created|updated)_at")), ~ sub("\\..*", "", .)) %>% 
  mutate_at(vars(matches("(created|updated)_at")), ymd_hms)
```

Let's just work with a random sample of 5% of the original dataset (`r as.numeric(nrow(ads_sample)/0.05)` rows).

## First-pass data summary

Looking at the volume of ads and impressions over time, regardless of whether it's marked as political by some indicator variable.

```{r ad_vol_basic, fig.asp=0.5}
ads_sample %>%
  group_by(wk = round_date(created_at, "1 week")) %>% 
  summarise(ads = n()) %>% 
  ggplot(aes(wk, ads)) +
  geom_bar(stat = "identity",
           fill = dutchmasters$little_street[[5]],
           alpha = 1/1.2) +
  geom_smooth(method = "loess", span = 0.6,
              colour = dutchmasters$little_street[[1]],
              se = FALSE, size = 1.1) +
  scale_y_continuous(breaks = pretty_breaks(4)) +
  labs(x = NULL, y = "# of Ads",
       title = "Weekly Ad Volume")
```

```{r impressions_vol_basic, fig.asp=0.56}
ads_sample %>%
  group_by(wk = round_date(created_at, "1 week")) %>% 
  summarise(impressions = sum(impressions)) %>% 
  ggplot(aes(wk, impressions)) +
  geom_bar(stat = "identity",
           fill = dutchmasters$little_street[[5]],
           alpha = 1/1.2) +
  geom_smooth(method = "loess", span = 0.6,
              colour = dutchmasters$little_street[[1]],
              se = FALSE, size = 1.1) +
  scale_y_log10(breaks = c(10,100,1e3,1e4)) +
  labs(x = NULL, y = "# of Impressions",
       title = "Weekly Impression Volume",
       subtitle = "log scale",
       caption = "Source: ProPublica Facebook Political Ad Collector")
```

And we can look at the weekly distribution of political ratings (based on the ProPublica classifier)

```{r political_frac}
ads_sample %>% 
  mutate(wk = round_date(created_at, "week")) %>% 
  ggplot(aes(created_at,
             political_probability)) +
  geom_boxplot(aes(group = wk),
               fill = dutchmasters$little_street[[5]],
               colour = dutchmasters$little_street[[6]],
               outlier.shape = NA) +
  geom_smooth(method = "loess", span = 0.5,
              colour = dutchmasters$little_street[[1]],
              se = FALSE, size = 1.1) +
  labs(x = NULL, y = NULL,
       title = "Political Probability of Ads",
       subtitle = "as classified by ProPublica Facebook Ad Collector",
       caption = "*dataset excludes ads with probability < 70%") +
  scale_y_continuous(labels = percent)
```

## Future work

- Highlighting weekly summary with notable events (i.e. special elections, particular ad buyers)
- Relationship between political probability and ad / impression volume
- Targets, advertisers, paid for by (and their relationships in time, with probability/vol/etc)

## In the weeds...

#### Numeric columns

```{r dists}
quiet_dist <- quietly(function(.x) distributions(.x, charts = FALSE))

ads_sample %>% 
  select_if(is.numeric) %>% 
  quiet_dist() %$% 
  DT::datatable(result, options = list(dom = "ft"))
```

#### Anomalies

```{r anomalies}
anomalies(ads_sample, 0.85, 1)[["variables"]] %>% 
  DT::datatable(extensions = c("Buttons", "Responsive"),
                options = list(dom = "Bft", buttons = I("colvis")))
```